---
title: "Giving `autograph` hints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Giving `autograph` hints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
if(Sys.info()[["sysname"]] == "Darwin")
  Sys.setenv(KMP_DUPLICATE_LIB_OK=TRUE)
  
reticulate::use_virtualenv("tf2")
```

```{r setup}
library(tensorflow)
library(keras)

library(tfautograph)
tf$version$VERSION
```

`autograph()` works without needing to provide any hints. It can automatically
translate expressions like `if` and `while` into `tf.cond()` and
`tf.while_loop()` calls. 

However, compared to the flexibility of R control flow, the tensorflow control
flow functions have much narrower constraints on how the control flow nodes are
constructed. The constraints of functions like `tf.cond()`, `tf.while_loop()`
and `dataset.reduct()` require that `autograph()` do some sophisticated
inference from the R code to anticipate what the outcome will be from evaluating
a block of code, so that it can "fix-up" the outcome to satisfy the constraints
tensorflow sets out.

If you want to bypass these inference routines, (for example, because the
inferred outcome is producing a graph that has more nodes than strictly
necessary, or for some minor performance benefits), you can provide hints to
`autograph` to let it know what it would otherwise try to infer.

There are 2 main types of hints, corresponding to the two main control flow
functions: `if` and `while`.

Since these hints are only relevant in graph mode we'll disable eager execution
for the rest of the vignette. In your code you're more likely to enter graph
mode only when tracing a `tf.function()`.
```{r}
tf$compat$v1$disable_eager_execution()
sess <- tf$compat$v1$Session()
```



# Passing hints to `if` 

Take the following expression:
```{r}
x <- tf$compat$v1$placeholder('float32')
y <- tf$constant(0)
```


```{r}
y
autograph({
  if(x > 0)
    y <- y + 1
})
y
```

Notice that `y` points to a different tensor after the `autograph()` call. The
new tensor has `'cond'` in the name, which is a big hint that `autograph()` took
that expression and called `tf.cond()` with it.

If you were to inspect the graph (e.g, by reading the `GraphDef` or inspecting
it in tensorboard, perhaps by calling `tfautograph::view_function_graph()`), you
would see that in essence, the expression got translated into the equivalent of:
```{r}
y <- tf$cond(x > 0, 
             true_fn = function() y + 1,
             false_fn = function() y)
```

`tf.cond()` takes the true and false branches as functions `true_fn` and
`false_fn`, and requires that the two functions return the same structure of
tensors.

How did `autograph()` know how to balance the two branches? It traced each
branch as a `tf.function` and then reconciled the differences. Then within the
`tf.cond()` call it recalled the traced function of the branch, along with some
routines that "fix-up" the actual outcome from the branch into a balanced final
outcome.

If you wanted to avoid tracing each branch as a separate `ConcreteFunction` and
balancing the output structures, you could supply a hint to autograph about what
outcomes you expect from the branches by calling `ag_if_vars()` right before the
`if`.

```{r}
autograph({
  ag_if_vars(y)
  if(x > 0)
    y <- y + 1
})
```

Here, `y` is the only variable modified, so it's the only symbol that we want
returned from the `tf.cond()`.

`ag_if_vars()` can help you specify a few different types of outcomes that
should be captured by the `tf.cond()` besides modified symbols, including
modified values in nested structures, return values, number of control flow
events (e.g., `break` and/or `next` statements), and local variables.
```{r}
args(ag_if_vars)
```


see `ag_if_vars()` for details.

# loops: `while` and `for`

Take the expression:
```{r}
x <- tf$constant(0)
n <- tf$constant(10)
autograph({
  while(n > 0) {
    x <- x + 1
    n <- n - 1
  }
})
```

If you were to build the equivalent with `tf.while_loop()` call it would look like this:
```{r}
c(x, n) %<-% tf$while_loop(
  cond = function(x, n)
    n > 0,
  body = function(x, n) {
    x <- x + 1
    n <- n - 1
    tuple(x, n)
  },
  loop_vars = tuple(x, n)
)
```

So, as you can see there are a few things that `autograph` needs to know to be
able to translate an R `while` loop into a `tf.while_loop()` call. Namely, it
needs to know what are the `loop_vars` that might be modified by the loop body,
both so that it can pass them to the `loop_vars` argument and so that it can
construct `cond` and `body` functions with the appropriate formals and return
values.

How does `autograph()` know which symbols belong in `loop_vars`, which are meant
to be local-only variables, and which are only read and never modified? There
are a couple different ways to approach solving this, but the way `autograph()`
does it by statically inferring them by examining the R loop body expression.
That is, without actually evaluating the code, we inspect which symbols are the
targets of common assignment operators like `<-`, `->`, `=`, `%<-%`, `%->%` and
`%<>%`. If you are autographing a `for` loop like `for(var in seq)` than the
symbol `var` is included in the set of symbols that would be statically inferred
as assignment targets.

After all the assigned symbols are inferred, they are classified as either
belonging to `loop_vars` or `undefs` based on whether the symbol already exists.


This approach is quicker and cheaper than tracing, but it can sometimes lead to
some symbols to `loop_vars` that otherwise could have been to be local-only
variables.


You can provide hints to autograph about what are the `loop_vars` like so:
```{r}
autograph({
  ag_loop_vars(x, n)
  while(n > 0) {
    x <- x + 1
    n <- n - 1
  }
})
```

Because the most common motivation for providing a loop hint will be to exclude
a certain symbol from being captured as a loop var, there is special syntax for
that. Prefix the variable with a minus `-` like so:
```{r, eval=FALSE}
tmp <- tf$constant(99)
autograph({
  ag_loop_vars(-tmp)
  while(n > 0) {
    tmp <- x + 1
    x <- tmp
    n <- n - 1
  }
})
```

Note that by default `ag_loop_vars` exports excluded variables as `undefs`,
symbols that throw an error (with a helpful error message) if you try to access
them. To prevent `autograph` from creating undefs you can pass NULL like so:

```{r, eval = FALSE}
ag_loop_vars(-tmp, undefs = NULL)
```

You can also use `ag_loop_vars()` to specify additional symbols to be included
with the statically inferred ones by prefixing them with a `+`. See
`?ag_loop_vars()` for additional examples.
