---
title: "Autographing Tensorflow version 1.15"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Autographing Tensorflow version 1.15}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(knitr)
.hook_error <- knit_hooks$get('error')
knit_hooks$set(error = function(x, options) {
  x <- strsplit(x, "[<", fixed = TRUE)
  x <- lapply(x, function(.) paste0("# [<", .))
  x <- unlist(x, FALSE, FALSE)
  .hook_error(x, options)
})

if(Sys.info()[["sysname"]] == "Darwin")
  Sys.setenv(KMP_DUPLICATE_LIB_OK=TRUE)
```

```{r setup}
Sys.setenv("RETICULATE_PYTHON" = reticulate::virtualenv_python("tf-1.15"))
library(tensorflow)
library(tfautograph)

tf$version$VERSION
tf <- tf$compat$v1
```

`autograph()` has complete compatibility with both Tensorflow version >= 2.0 and >= 1.15. For the most part, there are no difference between how `autograph()` work in both version, either in eager mode or graph mode. 

One thing to be aware of however is how control dependencies are registered. In Tensorflow version 2, all ops created inside a traced `tf.function()` are executed sequentially. That means that simply inlining a call to `tf$print()` or `tf$Assert()` is sufficient to ensure that the op is executed, and is executed in the intended order in a sequence of operations.

However, when tensorflow is not executing eagerly and outside a `tf.function()` context (that is, the default behavior of Tensorflow version 1, but also possible in version 2 after a call to `tf.compat.v1.disable_eager_execution()`), then created ops are only typically only evaluated if they are manually registered as control dependencies of other tensors.

For example, this block of code never throws an error:
```{r}
sess <- tf$Session()
x <- tf$constant(-1)

stop_if_negative <- function(x) {
  tf_assert(x > 0, "x is positive")
  x
}
sess$run(stop_if_negative(x))
```

There are two straight forward solutions to this. One is to wrap the block of code in an `tf.function()`
```{r, error = TRUE}
stop_if_negative2 <- tf_function(stop_if_negative)
sess$run(stop_if_negative2(x))
```


The other is to capture the assert operation as a control dependency of the evaluated tensor.
```{r, error = TRUE}
stop_if_negative3 <- function(x) {
  assert_op <- tf_assert(x > 0, "x is positive")
  with(tf$control_dependencies(list(assert_op)), {
    y <- tf$identity(x)
  })
  y
}

sess$run(stop_if_negative3(x))
```

Because the latter approach in `stop_if_negative3` is so common and so cumbersome, `autograph` takes care of it for you when autographing `stopifnot()` calls.

```{r, error = TRUE}
stop_if_negative4 <- function(x) {
  autograph(stopifnot(x > 0))
  x
}
sess$run(stop_if_negative4(x))
```
Note that the error was thrown even though we are not using `tf.function()`. How does `autograph(stopifnot(...))` guarantee that the created `tf$Assert()` Op is evaluated? With a multi-pronged, belt-and-suspenders approach:

+  it calls `tf$identity()` on every symbol involved in the assert expression and then replaces those symbols in the current scope.
+  it leaves a `tf$control_dependency()` context open for the rest of the current execution scope.
+  it registers a frame exit hook with `on.exit()` that calls `tf$identity()` on the return value and then closes the control dependency context.

`autograph(stopifnot(...))` only does these additional things outside of a `tf$function()` and when not executing eagerly. In side a `tf.function()` or when executing eagerly, `stopifnot()` merely calls `tf$Assert()`, and doesn't attempt to manually open and capture the ops as control dependencies.
